---
title: "Preliminary Run-through"
output: html_notebook
---

This notebook serves as a tester for a preliminary run-through of data cleaning for the weather data used in the OPUS grant project. I'll be following Joel's instructions on this script and getting an output that may be useful in future analyses. 

MET files are weather (datalogger) data associated with each mark-recapture study at Corfu WA.
Data fields in each .csv file:
"    "  “ID” "Date"   "Time"    "Sol"  "Ta"   "Tbe" "Tbs" "Tbw" . 

Necessary libraries:

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(weathermetrics)
```

For each file (start with Met.1991_2.csv file)
Read in file

```{r}
met912 <- read.csv("~/Desktop/GitHub/OPUS/Data/Met.1991_2.csv")
```

Rename X1 (first field) to CaseNum

```{r}
met912 <- met912 %>%
  rename(CaseNum = X)
```

Rename Date to JDate  [Julian day of year]

```{r}
met912 <- met912 %>%
  rename(JDate = Date)
```

Delete ID field

```{r}
met912 <- subset(met912, select = -c(ID))
```

Add Year field (value = year, e.g. 1991)

```{r}
met912$Year <- 1991
```

Add Date field (MO-DA-YEAR or similar format) based on Year and JDate. (Note that this depends on leap year or not.)

```{r}
#see below - getting it from NOAA data
```

Change outliers to NA:
	Ta, Tbe, Tbs, Tbw: <-20 OR >60
	Sol:  <-5 OR > 140
	
```{r}
max(met912$Ta)
min(met912$Ta)
max(met912$Tbe)
min(met912$Tbe)
max(met912$Tbs)
min(met912$Tbs)
max(met912$Tbw)
min(met912$Tbw)

max(met912$Sol)
min(met912$Sol)

```
	
No need to write a loop for this right this second - none lie outside of those values. 

Potential loop format: 

```{r}
for(i in 1:length(met912$CaseNum)){
  if(met912[i, 5] < -20){
    met912[i, 5] <- NA
  } else if (met912[i,5] > 60){
    met912[i, 5] <- NA
  }
}
```

	
Add Tbmean field:  Tbmean = mean(Tbe, Tbs, Tbw)  [mean model body temperature]

```{r}
met912$Tbmean <- NA

for(i in 1:length(met912$CaseNum)){
  vals <- c(met912[i, 6], met912[i, 7], met912[i, 8])
  met912[i, 10] <- mean(vals)
}
```

Add Tdelta field: Tdelta = Tbmean – Ta  [difference between mean body and air temperature]

```{r}
met912$Tdelta <- met912$Tbmean - met912$Ta
```

Make quick plot of Tbmean, Ta and Tdelta vs CaseNum (multiple lines on a single plot) to inspect data.

```{r}
ggplot(aes(x = CaseNum), data = met912) +
  geom_line(aes(y = Tbmean), color = "#95D840") + 
  geom_line(aes(y = Ta), color = "#3F87BC") + 
  geom_line(aes(y = Tdelta), color = "#440D54")

            
```

Make summaries by=JDate of Sol, Tbmean, Ta and Tdelta and plot if useful.

(What exactly do you want me to summarize? Means, max, min?)

```{r}
Solsum <- met912 %>%
  group_by(JDate) %>%
    summarise(mean = mean(Sol), max = max(Sol), min = min(Sol))

Tbmeansum <- met912 %>%
  group_by(JDate) %>%
    summarise(mean = mean(Tbmean), max = max(Tbmean), min = min(Tbmean))

Tasum <- met912 %>%
  group_by(JDate) %>%
    summarise(mean = mean(Ta), max = max(Ta), min = min(Ta))

Tdeltasum <- met912 %>%
  group_by(JDate) %>%
    summarise(mean = mean(Tdelta), max = max(Tdelta), min = min(Tdelta))
```



The goal is to develop daily metrics of weather [Sol and Ta] and body temperature [Tbmean and Tdelta] to associate with daily recapture data (e.g. see Corfu.1991nb2.NB_RECAPS_ALL.csv).


## Add in weather station data to compare to model temperature

Load in the dataset

```{r}
wsd <- read.csv("~/Desktop/GitHub/OPUS/Data/NOAAdata91Othello.csv")
```

Keep relevant columns

```{r}
wsd <- wsd %>%
  select(DATE, TMAX, TMIN)
```

Convert date to julian day 

```{r}
wsd$DATE <- as.Date(wsd$DATE, "%m/%d/%y")

wsd$JDate <- yday(wsd$DATE)
```

Convert from Fahrenheit to Celsius 

```{r}
wsd$TMAX <- fahrenheit.to.celsius(wsd$TMAX)
wsd$TMIN <- fahrenheit.to.celsius(wsd$TMIN)
```

Rename columns

```{r}
wsd <- wsd %>%
  rename(wsmax = TMAX, wsmin = TMIN, date = DATE)
```


Combine datasets - combine min and max temps into the Ta summary table (to compare our air temp recordings to the temp recordings from the weather station)

```{r}
Taws <- merge(Tasum, wsd, by = "JDate")
```

Plot differences in temps for max and min

```{r}
ggplot(aes(x = wsmax, y = max), data = Taws) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  xlim(25,45) + 
  ylim(25,45)
```

```{r}
ggplot(aes(x = min, y = wsmin), data = Taws) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  xlim(0,20) + 
  ylim(0,20)
```

